Signatures of Foreground Sky in Power Spectra of Redshifted Neutral Hydrogen from the Epoch of Reionization

### 2014-09-11 foreground_delay_spectrum_modeling.tex
From v0.5.1: ===========
(*) Mentioned cosmological parameters used in the paper.
(*) Wrote expression for relationship between k-parallel and eta
(*) Mentioned center frequency corresponds to redshift of 6.68
(*) Changed figure captions.

Since we are studying a redshifted H{\sc i} spectral line, $f$ is a measure of cosmological distance along the line of sight. Consequently, $\eta$ is a measure of line of sight spatial frequency mode, $k_\parallel$. Thus, $|V_\eta(\overline{\mathbf{u}},\eta)|^2$ is directly related to the spatial power spectrum of redshifted H{\sc i} distribution, $|V(\overline{\mathbf{k}}_\perp,k_\parallel)|^2$. However, foreground objects emit in radio frequencies and contaminate the signals from the redshifted H{\sc i}. Due to their spatial and spectral properties, $V_\eta(\overline{\mathbf{u}},\eta)$ also contains contamination from foreground radio objects \citep{thy13,tro12,mor12,bow09}. These contaminations have to be characterized precisely in order to reduce their impact on redshifted H{\sc i} power spectrum detection sensitivity.

@@

@@

The MWA was completed in 2012, and after commissioning, began its EoR observing program in 2013. In its final configuration the MWA is a 128--tile interferometer array capable of observing a 30~MHz band anywhere in the range 80--300~MHz. Each tile is a phased array of 16 dipoles each in the shape of a bow--tie characterized by a grating response with a $\gtrsim$~20\arcdeg wide primary field of view and multiple secondary side lobes. The tile arrangement is a centrally condensed core of $\sim$~300~m --- there are many spacings in the range 5--50~m --- and a radial density that falls off as the inverse of the radius with the longest baselines at 3~km \citep{bea12}. Here we focus on antenna spacings $|\overline{\mathbf{x}}| \le 200$~m (spatial scales relevant to reionization) and a frequency range 170--200~MHz.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
\caption{ Delay spectra of two 112s  Murchison Widefield Array observations of the Epoch of Reionization field at RA=0, Dec=-30. At top an {\it off--zenith}  pointing when tiles have significant side lobe response in the direction of the galactic center. At bottom a {\it zenith} pointing where side lobes are minimal and the galaxy has completely set. To get the the power spectrum each baseline is Fourier transformed into delay space ($y$--axis) and the amplitude plotted as a logarithmic color in order of baseline length ($x$--axis). The logarithmic color scale (shown at the right) is common to both panels. White lines mark the boundaries of the {\it foreground wedge} determined by the light travel time across each baseline. Smooth spectrum objects are physically limited to the region below this line. The various features within the wedge depend on the \emph{orientation} of the baseline, the side lobe structure of the power--pattern and the sky at that instant. \label{fig:fhd_data}}
\end{figure}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_sim_delay_spectrum_snapshots.eps}
\caption{Delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with simulations for {\it off--zenith} (top) and {\it zenith} (bottom) pointings. White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing. The features resemble those obtained with MWA data shown in figure~\ref{fig:fhd_data}. The Galactic center and Galactic plane prominently visible in the {\it off--zenith} pointing makes the emission in the {\it foreground wedge} brighter relative to that in the {\it zenith} pointing. The axes and color scale are identical to those in figure~\ref{fig:fhd_data}. \label{fig:sim_data}}
\end{figure}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_dsm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
\caption{Delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with simulations for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse foreground model without any thermal noise. White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing. The logarithmic color scale is common to both panels. In the {\it zenith} pointing, diffuse emission delay spectrum has a {\it two--pronged fork}--shaped structure and is present even at wide antenna spacings. \label{fig:noiseless-dsm-delay-spectrum}}
\end{figure}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_csm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
\caption{Delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with simulations for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the compact objects foreground model without any thermal noise. White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing. The logarithmic color scale is identical to that in Figure~\ref{fig:noiseless-dsm-delay-spectrum}. Emission from compact objects gives rise to centrally concentrated features in the {\it foreground wedge}.\label{fig:noiseless-csm-delay-spectrum}}
\end{figure}

>>>>

Since we are studying a redshifted H{\sc i} spectral line, $f$ is a measure of cosmological distance along the line of sight. Consequently, $\eta$ is a measure of line of sight spatial frequency mode, $k_\parallel$, given by $k_\parallel\approx 2\pi\eta\,f_{21}H_0\,E(z)/(c(1+z)^2)$, where $f_{21}$ is the rest frame frequency of 21~cm spin flip transition of H{\sc i}, and $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology. Thus, $|V_\eta(\overline{\mathbf{u}},\eta)|^2$ is directly related to the spatial power spectrum of redshifted H{\sc i} distribution, $|V(\overline{\mathbf{k}}_\perp,k_\parallel)|^2$. However, foreground objects emit in radio frequencies and contaminate the signals from the redshifted H{\sc i}. Due to their spatial and spectral properties, $V_\eta(\overline{\mathbf{u}},\eta)$ also contains contamination from foreground radio objects \citep{thy13,tro12,mor12,bow09}. These contaminations have to be characterized precisely in order to reduce their impact on redshifted H{\sc i} power spectrum detection sensitivity.

@@

Wherever applicable, we use $H_0=70$~km~s$^{-1}$~Mpc$^{-1}$ and assume a flat $\Lambda$CDM universe with $\Omega_\textrm{M}=0.27$ and $\Omega_\Lambda=0.73$.

@@

The MWA was completed in 2012, and after commissioning, began its EoR observing program in 2013. In its final configuration the MWA is a 128--tile interferometer array capable of observing a 30~MHz band anywhere in the range 80--300~MHz. Each tile is a phased array of 16 dipoles each in the shape of a bow--tie characterized by a grating response with a $\gtrsim$~20\arcdeg wide primary field of view and multiple secondary side lobes. The tile arrangement is a centrally condensed core of $\sim$~300~m --- there are many spacings in the range 5--50~m --- and a radial density that falls off as the inverse of the radius with the longest baselines at 3~km \citep{bea12}. Here we focus on antenna spacings $|\overline{\mathbf{x}}| \le 200$~m (spatial scales relevant to reionization) and a frequency range 170--200~MHz. The center frequency 185~MHz corresponds to a redshift, $z\simeq 6.68$.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
\caption{ Delay spectra of two 112s  Murchison Widefield Array observations of the Epoch of Reionization field at RA=0, Dec=-30. At top an {\it off--zenith}  pointing when tiles have significant side lobe response in the direction of the galactic center. At bottom a {\it zenith} pointing where side lobes are minimal and the galaxy has completely set. To get the the power spectrum each baseline is Fourier transformed into delay space ($y$--axis) and the amplitude plotted as a logarithmic color in order of baseline length ($x$--axis). For reference, the $k$--space equivalent of $\eta$ and $|\overline{\mathbf{x}}|$, namely, $k_\parallel$ and $k_\perp$ are shown at the axes on the right and top respectively. The logarithmic color scale (shown at the right) is common to both panels. White lines mark the boundaries of the {\it foreground wedge} determined by the light travel time across each baseline. Smooth spectrum objects are physically limited to the region below this line. The various features within the wedge depend on the \emph{orientation} of the baseline, the side lobe structure of the power--pattern and the sky at that instant. \label{fig:fhd_data}}
\end{figure}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_sim_delay_spectrum_snapshots.eps}
\caption{Delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with simulations for {\it off--zenith} (top) and {\it zenith} (bottom) pointings. White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing. The axes and color scale are identical to those in figure~\ref{fig:fhd_data}. The features resemble those obtained with MWA data shown in figure~\ref{fig:fhd_data}. The Galactic center and Galactic plane prominently visible in the {\it off--zenith} pointing makes the emission in the {\it foreground wedge} brighter relative to that in the {\it zenith} pointing. \label{fig:sim_data}}
\end{figure}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_dsm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
\caption{Delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with simulations for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse foreground model without any thermal noise. White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing. The axes and color scale are identical to those in figures~\ref{fig:fhd_data} and \ref{fig:sim_data}. In the {\it zenith} pointing, delay spectrum from diffuse emission has a {\it two--pronged fork}--shaped structure and is present even at wide antenna spacings. \label{fig:noiseless-dsm-delay-spectrum}}
\end{figure}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_csm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
\caption{Delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with simulations for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the compact objects foreground model without any thermal noise. White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing. The axes and logarithmic color scale are identical to those in figures~\ref{fig:fhd_data}, \ref{fig:sim_data} and \ref{fig:noiseless-dsm-delay-spectrum}. Emission from compact objects gives rise to centrally concentrated features in the {\it foreground wedge}.\label{fig:noiseless-csm-delay-spectrum}}
\end{figure}
